Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-06-04 paper/main.tex
Auto Update from local Sharelatex

This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of published privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information though synthetic means or anonymization. We collectively refer to de-identified data and synthetic data as \emph{faux data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but wihout the problematic derivation from live animals and reliance upon limited resources.

@@

% some text about goal to make our research reproducible/replicable - provide source code, pip/conda environment files, docker container
% explain why not full synthesize a full dataset. better to limit to sql query where relevant data is brought together; easier to model 1 table than lots of separate tables. Modelling lots of separate tables requires extensive meta data configuration and a significant amount of computing power to generate...
% where to put this text? "Unit testing is important but often overlooked - unit tests can show how a researcher validated their code was giving expected results which can often lead to insights not normally communicated in a results section of a journal article."

% Logical step our work is taking. 
One obvious way around this pickle is to separate data and analysis in such a way that the analysis can be performed on data without the \emph{analyst} having access to the data. This is fairly straight-forward to implement. Essentially, the analyst, with only a description of the sensitive data, composes analytical code which is handed off to the data custodian. The custodian then runs it and returns the results to the researcher.

This has been technologically possible for some time, but it's awkward in practice because activities such as data cleaning, exploration, and plotting are interactive and iterative processes which would become impractical with a mediator.

% What is our major contribution to the area of reproducibility in machine learning?
Kung Faux Pandas sidesteps this problem by generating \emph{faux-data} that is ``statistically similar'' to the real data but does not itself contain any sensitive data.

% this is a key unique feature of this work and should likely be repeated in conclusion. Perhaps also include in abstract
This faux-data can be prodded, poked, plotted, and posted for the world to see, enabling analysts to develop their code using whatever means they prefer. When the analytical code is ready, it can then be sent to the data custodian to run on the real data.

Many methods of generating faux-data exist in the literature, many of which have software implementations freely available. KFP provides a standard mechanism for ``wrapping'' any method into a ``plug-in'' which integrates the method with the Pandas data frame model. Three of these plug-ins are provided with full documentation for creating other plug-ins.

@@

The ideal environment for a machine learning developer to develop, test, or validate a model is to have unfettered access to all data in the problem domain they are working in. However, as discussed previously, legal limitations and data protection processes are slow. Figure~\ref{fig:architecture} depicts a basic architecture of an alternative to direct data access: data access through an intermediary that can synthesize data on-demand.

@@

Figure~\ref{fig:synthesis_process} provides a detailed look at the automated faux data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, machine learning developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.

% somewhere - here? Explain that the environment has been provided with a randomly generated dataset similar to one researchers have worked with during anlysis of a Sepsis prediction model.

@@

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a machine learning developer can access on-demand faux data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a machine learning process. Alternatively, an http REST service (utilized by the SPA) allows for cross language compatible requesting and receiving of faux data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible machine learning. Through providing ad-hoc faux data, machine learning researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing minimum dataset size issues, and evaluation of this system in a machine learning workflow.

While we have provided software code and a description of our methods in our GitHub repository, we believe that computational reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in machine learning research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

@@

>>>>

This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of published privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information though synthetic means or anonymization. We collectively refer to de-identified data and synthetic data as \emph{faux-data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but without the problematic derivation from live animals and reliance upon limited resources.

@@

% where to put this text? "Unit testing is important but often overlooked - unit tests can show how a researcher validated their code was giving expected results which can often lead to insights not normally communicated in a results section of a journal article."
One way around the difficulties of source data access is to separate the dependency between data and analysis in such a way that analysis can be performed without the analyst having access to the source data. Such a process can be possible if a sufficiently detailed description of the sensitive data is provided. Appropriate models can be developed and initially validated before either being handed off to a data custodian who then runs the processes on protected data and returns the results to the researcher or access to the protected data has been granted directly to the researcher. Solutions of this form have been possible for some time, but are awkward in practice because activities such as data cleaning, exploration, and model building are interactive and iterative processes that require more than just descriptions of data.

KFP addresses the issue of data access by generating faux-data that is ``statistically similar'' to the real data but does not itself contain any sensitive data. This faux-data can be analyzed, cleaned, feature-engineered, etc. as if it were the actual source data. KFP provides a standard mechanism for wrapping a privacy protection method into a plug-in which integrates the method with the Pandas DataFrame model. Three of these plug-ins are provided with full documentation for creating other plug-ins.

As the data generated through KFP has no protected component, it can be shared along with code or methods to improve reproducibility as well as aid in replicability. To further the aim of reproducibility we have provided our code and a sample of a randomly generated data set for testing purposes through our GitHub repository in a range of formats: Python environment files (conda and pip) to allow others to start with the same software libraries and specific versions that we used, a Docker container with code and sample data installed and already configured, and all source code along with version history so others can follow our development process.

KFP focuses on generating faux-data narrowed to a specific subset that a researcher is interested in rather than attempting to generate an entire population of data such as an entire EHR or longitudinal patient record. By focusing on just the data needed, performance is improved, the need for extensive modeling or metadata configuration is reduced, and the amount of faux-data generated and stored is minimized. Although the plugable methods included for faux-data generation have been shown to effectively replace source data for analysis, KFP facilitates a more through process that includes an initial exploration with faux-data followed by a repetition of methods on original source data through query logging and thereby aiding in replication.

@@

The ideal environment for a machine learning developer to develop, test, or validate a model is to have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} depicts a basic architecture of an alternative to direct data access: data access through an intermediary that can synthesize data on-demand.

@@

Figure~\ref{fig:synthesis_process} provides a detailed look at the automated faux-data request/response process. All queries submitted to the system are first pre-processed to remove clauses such as "ORDER BY" which are made ineffective by the synthesis process. All queries are logged for later analysis and results cached to optionally improve synthesis performance for repeated queries. Next the processed query is executed against the protected data set and results are passed into the modular synthesis process. While the current implementation only utilizes a couple of synthesis methods, machine learning developers or data warehouse owners can insert their own custom method or modify existing methods. Previously removed "ORDER BY" clauses are re-applied to the faux-data before results are returned to the user. Finally, although not implemented in this version of KFP, synthetic results could be held back until reviewed and approved for release by the data owner.

@@

\begin{wrapfigure}[15]{r}{0.5\textwidth}%optionally place [15] after {wrapfigure} to control how many lines to wrap around figure
  \centering
  \includegraphics[width=0.5\textwidth]{ui_screenshot3}
  \caption{User Interface.}
  \label{fig:ui}
\end{wrapfigure}
KFP provides multiple methods by which a machine learning developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a machine learning process. Alternatively, an http REST service (utilized by the SPA) allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible machine learning. Through providing ad-hoc faux-data, machine learning researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. However, there are several areas that could be improved or researched further: improved code sharing, addressing minimum dataset size issues, and evaluation of this system in a machine learning workflow.

While we have provided software code and a description of our methods, we believe that computational reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in machine learning research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

@@

Rather than the big step of full EHR synthesis or anonymization, KFP takes a small step in being a more usable small data generation tool. Big steps are harder to adopt...
